Effects of Reynolds Number on the Performance and Near-Wake of a Vertical-Axis Cross-Flow Turbine

### 2015-04-18 paper.tex
Add to foil performance section in the introduction

Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
which is a function of the foil's angle of attack $\alpha$. The angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
in general is caused by the blade boundary becoming more turbulent as a fraction
of the chord length, diminishes the negative effects of a laminar separation
bubble occurring near the leading edge. A review of low Reynolds number airfoil
behavior is presented in \cite{Lissaman1983}.

As previously mentioned, the near-wake of a CFT is not very similar to a simple
drag source, or plane shear flow. One of the mean wake's streamwise vortex pairs
can be predicted by potential flow theory as a result of finite-span blades
producing lift (blade end effects), but the other---caused by viscous effects,
i.e., dynamic stall---is not. We can attempt to predict changes
\todo[inline]{Finish this thought with Reynolds number dependence of these
simpler flows, and set up the question whether or not we can assume these will
combine linearly to produce the wake $Re$-dependence we see.}

\begin{figure}[ht]
\caption{Airfoil lift to drag ratio as a function of Reynolds number, from 
\cite{McMasters1980}.}
\end{figure}

@@

Static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies the occurrence of dynamic stall at tip speed ratios at and below
those of maximum rotor torque\cite{Para2002}. We then must understand the
effects of Reynolds number on the dynamic stall process.

@@

@@

We expect that lower lift on the blades at lower $Re$ will weaken tip vortex
shedding, and decrease the levels of turbulence shed from the blade boundary
layers. As mentioned previously, the dynamic stall vortex shedding is not
expected to have a large Reynolds number sensitivity.

>>>>

Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
which is a function of the foil's angle of attack $\alpha$. The angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
in general is caused by the blade boundary becoming more turbulent as a fraction
of the chord length, diminishes the negative effects of a laminar separation
bubble occurring near the leading edge. For smooth airfoils, this transition can
cause a dramatic increase in foil performance, a shown in
Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
presented in \cite{Lissaman1983}.

\begin{figure}[ht]
\caption{Airfoil lift to drag ratio as a function of Reynolds number, from 
\cite{McMasters1980}.}
\label{fig:McMasters}
\end{figure}

@@

Static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies dynamic loading. Furthermore, the blades commonly undergo
dynamic stall for tip speed ratios at and below those of maximum rotor
torque\cite{Para2002}. We then must understand the effects of Reynolds number on
the dynamic stall process.

@@

\todo[inline]{Find reference to describe $Re$ effects on dynamic foil behavior.}

@@

We expect that lower lift on the blades at lower $Re$ will weaken tip vortex
shedding, and decrease the levels of turbulence shed from the blade boundary
layers. As mentioned previously, the dynamic stall vortex shedding is not
expected to have a large Reynolds number sensitivity, though a larger separation
bubble at lower $Re$ may induce higher levels of turbulence as shed vortices
become unstable and break down.
